---
title: "Visualization"
author: "Sameera"
date: "2026-10-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

Scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y= tmax))
```

```{r}
ggplot(weather_df, aes(x = tmin, y= tmax)) +
  geom_point()
```

Another way to do it where u put the dataframe first and 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Making it fancier

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point(alpha = .25) ##for blending
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, colour = name)) +
  geom_point(alpha = .25) + ##for blending
  geom_smooth(se = FALSE) ## gives smooth curve and se false removes standard error shading
```

where u defind your aesthetic matters

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(colour = name), alpha = .25) + ##for blending
  geom_smooth(se = FALSE) ## gives smooth curve and se false removes standard error shading
```

Show faceting

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y= tmax, colour = name)) +
  geom_point(alpha = .5) +
  facet_grid(. ~ name) ##separates the diff graphs 
```
by column or row depending on where u put the dot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y= tmax, colour = name)) +
  geom_point(alpha = .5) +
  facet_grid(name ~ .) ##separates the diff graphs 
```

another way to do it

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y= tmax, colour = name)) +
  geom_point(alpha = .5) +
  facet_grid(cols = vars(name)) ##separates the diff graphs 
```

Moving on from temp

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, colour = name)) +
  geom_point(aes(size = prcp, alpha = .5)) + ## changes point size based on precipitation
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

plot only central park tmin and tmax and convert temp to fahrenheit

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  mutate(tminf = tmin * 1.8 + 32, 
         tmaxf = tmax * 1.8 + 32) |> 
  ggplot(aes(x = tminf, y = tmaxf)) +
  geom_point()
```

Hex plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Density plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax, colour = name)) +
  geom_density()
```
```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```

Boxplot

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

Violin plots

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

Ridge plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

## Save some plots

```{r}
ggp_weather =
weather_df |> 
  ggplot(aes(x = date, y = tmax, colour = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  facet_grid(. ~ name)

ggsave("images/ggp_weater.pdf", ggp_weather)
```

```{r fig.width = 12, fig.height = 12}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

























